---
title: "NYT_AD-Click_Analysis"
author: "Manjula Kottegoda"
date: "June 27, 2016"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Prep for Live Session 8
#### Module 8.9

Download the clickstream data from <http://stat.columbia.edu/~rachel/datasets/nyt13.csv>. And load to a Dataframe.

```{r}
# Load data to a Data frame
#NYT_DF13 <- read.csv("Analysis/Data/nyt13.csv",blank.lines.skip=TRUE)
#NYT_DF13 <- read.csv(url("http://stat.columbia.edu/~rachel/datasets/nyt1.csv"))
NYT_DF13 <- read.csv("Analysis/Data/nyt1.csv",blank.lines.skip=TRUE)
head(NYT_DF13) # Examine 1-5 records of Data frame
str(NYT_DF13) # Examine structure of Data frame
summary(NYT_DF13) # Examine summery of Data frame
```


###  Create a new variable ___ageGroup___ that categorizes age into following groups: 

####    < 18, 18–24, 25–34, 35–44, 45–54, 55–64 and 65+

```{r}
#Before cutting we look at the distribution of the Age values
hist(NYT_DF13$Age, main="Histogram of Age Distribution", xlab="Age")

#Add a new column called age group to Data frame and add designation based on the age intervals
NYT_DF13$ageGroup <- cut(NYT_DF13$Age, c(-Inf, 18, 24, 34, 44, 54, 64, Inf))

#Step above created values such as (-Inf,18]  (18,24]  (24,34]  (34,44]  (44,54]  (54,64]  (64, Inf]  
# this needs to be cleaned up to show ranges like below using the to the levels attribute of a variable
levels(NYT_DF13$ageGroup) <- c("<18", "18-24", "25-34", "35-44", "45-54", "55-64", "65+")

#Now we look at the data frame again
head(NYT_DF13) # Examine 1-5 records of Data frame with ageGroup column
str(NYT_DF13) # Examine structure of Data frame with ageGroup column

```

###For a single day (like May 14):
###Plot distributions of number impressions and click-through-rate (CTR = click/impression) for the age groups.

CTR is calculated by deviding clicks by Impressions. So get a copy of dataframe and drop rows with 0 values from it. Then Add another column wit the CTR value.

```{r}
NYT_DF13_CTR <- subset(NYT_DF13, Impressions>0)
NYT_DF13_CTR$CTR <- NYT_DF13_CTR$Clicks/NYT_DF13_CTR$Impressions
#Now we look at the new data frame 
head(NYT_DF13_CTR) # Examine 1-5 records
str(NYT_DF13_CTR) # Examine structure


#Now plotI Impressions
library(ggplot2) # used for visualizations
ggplot(subset(NYT_DF13_CTR, Impressions>0), aes(x=Impressions, fill=ageGroup))+
    geom_histogram(binwidth=1)


#Now plotI CTR
ggplot(subset(NYT_DF13_CTR, CTR>0), aes(x=CTR, fill=ageGroup))+
    labs(title="click-through-rate (CTR = click/impression) for the age groups for May 14")+
    geom_histogram(binwidth=.025)
```
